---
title: "DATA119 - Randomness and Simulation"
output: 
  learnr::tutorial:
    css: css/custom-styles.css
runtime: shiny_prerendered
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

library(learnr)
library(gradethis)
library(reticulate)

# Set the path to the existing Python environment
#reticulate::use_python("/opt/python/3.9.21/bin/python", required = TRUE)

# Optional: Install necessary Python packages if not already installed
#reticulate::py_install(c('numpy', 'pandas'))

custom_checker <- function(label, user_code, solution_code, envir_result, evaluate_result, last_value, stage, ...) {
  if (stage == "code_check") {
      if (user_code == solution_code){
          return(list(message = random_praise(), correct = TRUE))
      }
    return(list(message = random_encouragement(), correct = FALSE))
  }
}

gradethis_setup()
tutorial_options(exercise.completion = FALSE, exercise.checker = custom_checker)
```

```{r header, echo = FALSE}
library(htmltools)

tags$div(
  class = "topContainer",
  tags$div(
    class = "logoAndTitle",
    tags$img(
      src = "./images/dsi_logo.png",
      alt = "DSI Logo",
      class = "topLogo"
    ),
    tags$h1("Randomness and Simulation", class = "pageTitle")
  )
)
```

## Goals

The goals of this lab are:

* To learn how to use `np.random.choice` to select random values in python
* To practice using simulation to estimate probabilities
* To practice using simulations to improve decision making

## Randomness

intro here

### Overview

brief overview of lab and first setup chunk

### Introduction to `np.random.choice`

description then some practice exercises. similar to textbook

## Simulation

make an example 

do $1 milkshake exercise 

## Making the Best Lineup

take in stats from 2024 season for the Chicago White Sox
simulate different lineup combinations (mention caveats)
