---
title: "DATA119 - Data Visualization"
output: 
  learnr::tutorial:
    css: css/custom-styles.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

library(learnr)
library(gradethis)
library(reticulate)

# Set the path to the existing Python environment
#reticulate::use_python("/opt/python/3.9.21/bin/python", required = TRUE)

# Optional: Install necessary Python packages if not already installed
reticulate::py_install(c('numpy', 'pandas', 'plotnine', 'seaborn', 'altair', 'matplotlib'))

custom_checker <- function(label, user_code, solution_code, envir_result, evaluate_result, last_value, stage, ...) {
  if (stage == "code_check") {
      if (user_code == solution_code){
          return(list(message = random_praise(), correct = TRUE))
      }
    return(list(message = random_encouragement(), correct = FALSE))
  }
}

gradethis_setup()
tutorial_options(exercise.completion = FALSE, exercise.checker = custom_checker)
```

```{r header, echo = FALSE}
library(htmltools)

tags$div(
  class = "topContainer",
  tags$div(
    class = "logoAndTitle",
    tags$img(
      src = "./images/dsi_logo.png",
      alt = "DSI Logo",
      class = "topLogo"
    ),
    tags$h1("Data Visualization", class = "pageTitle")
  )
)
```

## Goals

The goals of this lab are:

* To understand the basic theory underpinning the field of data visualization.
* To learn which visualizations are best suited to which use cases
* To practice using matplotlib to make visualizations
* To get exposure to seaborn, plotnine, plotly, and altair as alternative graphing packages to matplotlib

## Visualization Theory

A clear graphic can convey the results of data science work more quickly and effectively than nearly any other approach. The way to design and implement the optimal graphic is the subject of continued research. Every element of a visualization, from the color palette to the graph type has an impact on its ability to communicate the intended information.

### Overview

In this lab you'll be introduced to two of the main frameworks for visualization used in Python. First, to the object-oriented approach used by `Matplotlib` and later the Grammar of Graphics approach used by libraries like `plotnine` and `altair`. We'll also briefly look at `seaborn` which is a hybrid of the two frameworks.

### Matplotlib Paradigm
Matplotlib takes an object-oriented approach that is akin to creating a painting. First you choose the size of your (virtual) canvas, then which specific points, lines, and polygons to draw on the canvas, and finally add finishing touches like titles, labels, legends, and styling. 
Common graph types like bar charts, line plots, and scatter plots have dedicated methods which take the data you provide as input, and produce the completed visualization. Though it may seem tedious, one advantage of Matplotlib is that it gives you fine-grained control for every element of your visualization. 
Since Matplotlib is one of the most commonly used packages for creating visualizations with Python, it will be the focus of this lab.

### Grammar of Graphics Paradigm
By contrast, the Grammar of Graphics approach is more like commissioning a painting from an artist. You simply tell it the data set to use, what type of graph you'd like, any particular coloring and scaling, and then the implementation (such as `plotnine` or `altair`) will take care of the rest. An advantage of Grammar of Graphics is the clarity that is enforced by its clear structure; however, you lose some of the fine-grained control available in object-oriented approaches like `Matplotlib`. 
This approach is based on the visualization theory developed by Leland Wilkerson, and implemented popularly in R with ggplot2 by Hadley Wickham.

## Setup

This lab will use `Python` and the package `pandas`. For graphing we'll primarily use `matplotlib`. We'll also see some `seaborn`, `altair`, and `plotnine`.
To explore these libraries, we'll be using a cleaned version of the dataset obtained from the [Chicago Open Data Portal](https://data.cityofchicago.org/Parks-Recreation/Beach-Weather-Stations-Automated-Sensors/k7hf-8y75/about_data), which contains data generated by three beach weather stations in Chicago. Our version of the data only contains records from 2024. Run this cell to load the packages and DataFrame.

```{python packages, exercise=TRUE, exercise.eval = FALSE, message = FALSE}
# data manipulation
import numpy as np
import pandas as pd
# graphing
import matplotlib.pyplot as plt
import altair
import seaborn as sns
import plotnine as p9

# so we can see all columns
pd.options.display.max_columns = 10

# To download this file go to ADD A HOSTED VERSION OF THIS FILE LATER

beach = pd.read_csv("./data/beach_data.csv", index_col = 0)
beach.head()
```


```{python main-setup, exercise=FALSE, include=FALSE, message=FALSE, exercise.setup="packages"}
beach_year_avg = beach.drop(columns=['Month', 'Hour', 'Day of Year']).groupby("Station Name").mean().reset_index()
stations = beach_year_avg['Station Name']
wind_avg = beach_year_avg['Wind Speed']
max_wind = beach[beach['Maximum Wind Speed'] < 100]['Maximum Wind Speed'] 
# dealing with weird outliers
hourly_wind = beach[beach['Maximum Wind Speed'] < 100]['Wind Speed']

beach_day_max = beach[beach['Station Name'] == 'Oak Street'][['Day of Year', 'Air Temperature']].groupby('Day of Year').max().reset_index().sort_values('Day of Year')

day = beach_day_max['Day of Year']
daily_temp = beach_day_max['Air Temperature']

beach_day_max_foster = beach[beach['Station Name'] == 'Foster'][['Day of Year', 'Air Temperature']].groupby('Day of Year').max().reset_index().sort_values('Day of Year')

daily_temp_oak = beach_day_max['Air Temperature']
daily_temp_foster = beach_day_max_foster['Air Temperature]
```

## Graphing with Matplotlib

Below you will be walked through the key components of Matplotlib. As you explore, it may be useful to check out [Matplotlib's quick start guide](https://matplotlib.org/stable/users/explain/quick_start.html) or other parts of its documentation for more details.

### Choosing your canvas: Figures and Axes
In Matplotlib the **figure** is the core element which sets the overall size of our canvas, and keeps track of everything being added to the canvas. Our **axes** are how we specify where on our figure we want our visualization to happen.   

Below you'll see a quick example of setting up the figure and axes before making a scatterplot. Feel free to experiment with the figure and axes by changing the `width`, `height`, `left_start`, `bottom_start`, `width_start`, and `height_start` variables.

```{python figure, exercise = TRUE, message = FALSE, exercise.eval=FALSE, exercise.setup="packages"}
width = 12
height = 6
left_start, bottom_start, width_start, height_start = (0.5, 0.5, .5, 0.5)
# do not edit below here
fig = plt.figure(figsize=(width, height))
ax = fig.add_axes((left_start, bottom_start, width_start, height_start), frame_on=True)
ax.scatter([1,2,3], [1,2,3])
plt.show()
```
Often we'll just let Matplotlib use the default figure with an axes spanning the entire figure size. Here's an example of that below using `subplots`.
```{python subplots, exercise = TRUE, message = FALSE,exercise.eval=FALSE, exercise.setup="packages"}
fig, ax = plt.subplots()
ax.scatter([1,2,3], [1,2,3])
plt.show()
```
Alternatively we can just create a plot without explicitly setting a figure or axes, though Matplotlib will be making them in the background automatically!
```{python no-figure, exercise = TRUE, message = FALSE, exercise.eval=FALSE, exercise.setup="packages"}
plt.scatter([1,2,3], [1,2,3]) 
# note how we have to do plt not ax.scatter
# this is because the axes hasn't been created when we call .scatter
plt.show()
```
Throughout the rest of this notebook we'll let Matplotlib handle making the figures and axes by default, but knowing they exist and how to manipulate them can be useful as you explore more complicated graphing techniques. 
Check out this [part of the documentation](https://matplotlib.org/stable/users/explain/axes/index.html) if you want to learn more about figures and axes!

### Common Graphs

As data scientists, it is important we choose the appropriate graph types for our data. 
As you practice making the different graphs below, think critically about which types of data each graph is best suited for. 

#### Bar Chart
A bar chart is useful for displaying summaries or distributions of categorical information. 
Category names are often written on the x-axis, with the height of each bar indicating the desired summary information for each category on the y-axis. 
In Matplotlib, the simplest way to make a bar chart is by using `plt.bar`. We provide a collection of labels as the first argument (`x`) and a collection of summary information as the second argument (`height`). 
Below is an example using a synthetic grade distribution.

```{python basic-bar, exercise = TRUE, message = FALSE, exercise.eval=FALSE, exercise.setup="packages"}
grade_categories = ["A", "B", "C", "D","F"]
students_per_category = [14, 10, 5, 2, 1]
plt.bar(x=grade_categories, height=students_per_category)
```

1. Assume we have stored the names of the Beach Weather stations in `stations` and their average wind reading for 2024 in `wind_avg`. 
Write one line of code to make a bar chart displaying the average wind recorded at each weather station in 2024.
```{python beach-avg-setup, include=FALSE, exercise = FALSE, message = FALSE, exercise.setup="packages"}
beach_year_avg = beach.drop(columns=['Month', 'Hour', 'Day of Year']).groupby("Station Name").mean().reset_index()
stations = beach_year_avg['Station Name']
wind_avg = beach_year_avg['Wind Speed']
```
```{python beach-avg-wind-bar, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-avg-setup"}

```
```{python beach-avg-wind-bar-solution, message = FALSE, warning = FALSE, echo = FALSE}
plt.bar(stations, wind_avg)
```

```{r beach-avg-wind-bar-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Great Bars",
)
```

Bar charts can help us quickly compare between different categories. Checkout the [Matplotlib Bar Chart Documentation](https://matplotlib.org/stable/api/_as_gen/matplotlib.pyplot.bar.html) to see more ways to customize your bar charts.

#### Histogram
Histograms are useful when we want to look at the distribution of 1-dimensional numerical data. By separating the data into bins, then plotting the count or percentage of data in each bin, we can visualize an entire distribution at a glance. 
In Matplotlib, the simplest way to make a histogram is with `plt.hist(x, bins, density)` where `x` is the data to be plotted, `bins` is the number of bins, and `density` changes the value shown on the y-axis. 
Play around with the bins and density arguments below to see how they effect the histogram. 

```{python basic-hist, exercise = TRUE, message = FALSE, exercise.eval=FALSE, exercise.setup="packages"}
np.random.seed(42) # for consistent samples
numbers = np.random.choice(np.arange(100), size=50) # a random sample of numbers
plt.hist(x=numbers, bins=10, density=False) # making the graph
```

```{r bins-impact, echo=FALSE}
question("2. As you increase the number of bins, what is the impact on the histogram? CHECK ALL THAT APPLY",
         answer("Nothing happens"),
         answer("The bins are wider"),
         answer("The bins are narrower", correct=TRUE),
         answer("There is less white space"),
         answer("There is more white space", correct=TRUE), 
        allow_retry = TRUE,
  random_answer_order = TRUE)
```
3. Make your own count histogram with 10 bins using the hourly maximum wind readings which have been stored for you in the variable `max_wind`. You can skip the density argument.
```{python beach-hist-setup, include=FALSE, exercise = FALSE, message = FALSE, exercise.setup="packages"}
max_wind = beach[beach['Maximum Wind Speed'] < 100]['Maximum Wind Speed'] # dealing with weird outliers
hourly_wind = beach[beach['Maximum Wind Speed'] < 100]['Wind Speed']

```
```{python beach-max-wind-hist, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-hist-setup"}

```
```{python beach-max-wind-hist-solution, message = FALSE, warning = FALSE, echo = FALSE}
plt.hist(max_wind, bins=10)
```

```{r beach-max-wind-hist-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Histogram :)"
)
```

As you've seen histograms are a powerful tool to quickly summarize 1-D numerical data. Checkout the [Matplotlib histogram documentation](https://matplotlib.org/stable/api/_as_gen/matplotlib.pyplot.hist.html) to see more ways to fine tune your histograms.

#### Scatterplot
So far, we have looked at single dimensional data. However, it is often useful to look at the relationship between two dimensions (or variables) of our data. A scatterplot can help us quickly visualize that relationship, by plotting each `(x,y)` pair of values on a 2-D plane. 
In Matplotlib, the simplest way to make a scatterplot is by using `plt.scatter(x,y)` and providing a collection of data as the first argument (`x`) and a collection of different data for the same records in the same order as the second argument (`y`). 
Below is an example to see how it works, looking at two randomly generated samples of data.

```{python basic-scatter, exercise = TRUE, message = FALSE, exercise.eval=FALSE, exercise.setup="packages"}
np.random.seed(42) # for consistency
# generating random samples
x = np.random.choice(np.arange(100), size=200)
y = np.random.choice(np.arange(100), size=200)
# drawing the scatterplot
plt.scatter(x=x, y=y)
```

4. We have stored the hourly wind speeds in `hourly_wind` and the maximum for each hour in `max_wind`. Make a scatterplot chart displaying the relationship between the hourly and maximum wind recorded at the weather stations in 2024. Let the x-axis contain the hourly wind readings.
```{python beach-wind-scatter, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-hist-setup"}

```

```{python beach-wind-scatter-solution, message = FALSE, warning = FALSE, echo = FALSE}
plt.scatter(hourly_wind, max_wind)
```

```{r beach-wind-scatter-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Great Points!",
)
```

As expected, these two variables have a positive correlation. 
Beyond seeing relationships between variables, scatterplots can be used for exploratory data analysis. There is an outlier in our plot above, which reveals a point which was erroneously collected. Checkout the [Matplotlib Scatterplot Documentation](https://matplotlib.org/stable/api/_as_gen/matplotlib.pyplot.scatter.html) to see more ways to customize your scatterplots (i.e., coloring, resizing, or changing the shape of your points).

#### Line Plot
Line plots are also used when we want to look at the relationship between two variables in our data. However, unlike scatterplots, line plots typically use the x-axis to convey some measure of sequential data, and therefore are often used to display relationships over time. 
A line plot is made up of points connected to their closest neighbor (sequentially) by a line segment. 
In Matplotlib, the simplest way to make a line plot is by using `plt.plot(x,y)` and providing a collection of data as the first argument and a collection of different data for the same records in the same order as the second argument. 
Note that you can also plot single points this way by making x and y both single numbers. 
Here is an example graphing the sine function.

```{python basic-line, exercise = TRUE, message = FALSE, exercise.eval=FALSE, exercise.setup="packages"}
x = np.arange(-4, 4, 0.2)
y = np.sin(x)
# drawing the line plot
plt.plot(x, y)
```


5. We have stored the daily temperature high readings for the Oak Street beach in `daily_temp` and the day of the year in `day`. Make a line chart displaying the daily temperature high readings at that beach throughout 2024.
```{python beach-temp-line-setup, include=FALSE, exercise = FALSE, message = FALSE, exercise.setup="packages"}
beach_day_max = beach[beach['Station Name'] == 'Oak Street'][['Day of Year', 'Air Temperature']].groupby('Day of Year').max().reset_index().sort_values('Day of Year')

day = beach_day_max['Day of Year']
daily_temp = beach_day_max['Air Temperature']
```

```{python beach-temp-line, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-temp-line-setup"}

```

```{python beach-temp-line-solution, message = FALSE, warning = FALSE, echo = FALSE}
plt.plot(day, daily_temp)
```

```{r beach-temp-line-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Great Points!",
)
```

Line plots can help us quickly identify temporal relationships (or the lack thereof) between different variables in a dataset. Checkout the [Matplotlib Line Plot Documentation](https://matplotlib.org/stable/api/_as_gen/matplotlib.pyplot.plot.html) to see more ways to customize your line plots (i.e. coloring, resizing, or changing the style of your lines).


### Finishing Touches: Labels, Legends, and Tick marks
So far we have focused on simply making the graphs; however, no visualization is complete without an informative title and labels. 
At times it can also be useful to customize your tick marks, add a legend, or make other changes around the graph itself to ensure a clear and informative visualization.

Below you'll practice how to add axis labels, a title, and a legend as well as how to modify tick marks. To see more customization options check out [this section of the quickstart guide](https://matplotlib.org/stable/users/explain/quick_start.html#styling-artists).

After we've made a figure manually, or one has been generated automatically for us through built-in functions,we can add an x-axis label with `plt.xlabel('text')`, a y-axis label with `plt.ylabel('text')`, and a title we can do `plt.title('text')`.

6. Add the title `Average Wind by Beach in 2024`, the y-axis label `Avg Wind`, and the x-axis label `Beach Name` to the bar chart we made earlier. Copy and paste the prior code here, and add to it.

```{python beach-avg-wind-bar-labeled, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="main-setup"}

```

```{python beach-avg-wind-bar-labeled-solution, message = FALSE, warning = FALSE, echo = FALSE}
plt.bar(stations, wind_avg)
plt.xlabel("Beach Name")
plt.ylabel("Avg Wind")
plt.title("Average Wind by Beach in 2024")
```

```{r beach-avg-wind-bar-labeled-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Great Bars",
)
```

Sometimes when we are plotting multiple data series, it can be useful to include a legend. 
To do so, we often have to add a label argument to each individual plot.

7. Edit the code below so that the first line plot is given the label 'Oak Street Beach' and the second 'Foster Beach' in a new legend located in the lower center of the graph. If you're stuck check out the [legend documentation](https://matplotlib.org/stable/api/_as_gen/matplotlib.axes.Axes.legend.html).

```{python beach-line-overlaid-setup, include=FALSE, exercise = FALSE, message = FALSE, exercise.setup="beach-temp-line-setup"}
beach_day_max_foster = beach[beach['Station Name'] == 'Foster'][['Day of Year', 'Air Temperature']].groupby('Day of Year').max().reset_index().sort_values('Day of Year')

day = beach_day_max['Day of Year']
daily_temp_oak = beach_day_max['Air Temperature']
daily_temp_foster = beach_day_max_foster['Air Temperature']
```

```{python beach-temp-overlaid-line, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-line-overlaid-setup"}
plt.plot(day, daily_temp_oak, label=...)
plt.plot(day, daily_temp_foster, label=...)
...
```
```{python beach-temp-overlaid-line-solution, message = FALSE, warning = FALSE, echo = FALSE}
plt.plot(day, daily_temp, label="Oak Street")
plt.plot(day, daily_temp_foster, label="Foster")
plt.legend(loc='lower center')
```
```{r beach-temp-overlaid-line-code-check, message = TRUE, warning = FALSE}
grade_this_code(
  correct = "What an absolute legend!",
)
```



## Building on Matplotlib: Pandas and Seaborn

Because making visualizations is such a key part of the data science workflow, `pandas` has methods for data frames that call the `matplotlib` package behind the scenes, simplifying the graphing process. 

All of the graphs we've discussed can be made from a dataframe using `df.plot.graph_name(**args)`. 
If you're not sure how to implement a particular graph check out the **Plotting** section of the [pandas cheatsheet](https://pandas.pydata.org/Pandas_Cheat_Sheet.pdf) or the [pandas plotting documentation](https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.plot.html). 
Below you'll recreate some of the earlier graphs to harness the power of combining pandas and matplotlib.

Before you start on the remaining exercises, here is an example to see how `pandas` implements a basic bar chart differently than `matplotlib`. 
Note the need to create a series or DataFrame first.

```{python pandas-basic-bar, exercise = TRUE, message = FALSE, exercise.eval=FALSE, exercise.setup="packages"}
grade_categories = ["A", "B", "C", "D","F"]
students_per_category = [14, 10, 5, 2, 1]
grades = pd.DataFrame({"Grades" : grade_categories, "# Students" : students_per_category})
grades.plot.bar("Grades", '# Students')
```

For data that is already stored in a list, it may be easier to just use `matplotlib` directly. However, using pandas also automatically added a x-axis label, a legend label, and allowed us to just reference columns from a dataframe directly, making for a much smoother process when our data is already in a dataframe. 

8. Recreate the bar chart displaying the average wind recorded at each weather station in 2024 using the dataframe created for you below. 
Make sure it has the same labels as Question 6. If you would like a challenge, do it all in one line with the x-axis labels rotated to 0 degrees for better visibility.


```{python beach-avg-pandas-setup, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="packages"}
beach_year_avg = beach.drop(columns=['Month', 'Hour', 'Day of Year']).groupby("Station Name").mean().reset_index()
```
```{python beach-avg-wind-bar-pandas, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-avg-pandas-setup"}




```
```{python beach-avg-wind-bar-pandas-solution, message = FALSE,warning = FALSE, echo = FALSE}
beach_year_avg.plot.bar("Station Name", "Wind Speed", xlabel="Beach Name", ylabel="Avg Wind", title="Average Wind by Beach in 2024", rot=0)
```

```{python beach-avg-wind-bar-pandas-hint-1, message = FALSE,warning = FALSE, echo = FALSE}
checkout the df.plot documentation to see how we can supply the axis labels and title directly
```

```{python beach-avg-wind-bar-pandas-hint-2, message = FALSE,warning = FALSE, echo = FALSE}
checkout the rot argument of df.plot.bar for help rotating the xticks labels
```

```{r beach-avg-wind-bar-pandas-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Great Bars",
)
```

Notice that though we needed to manipulate the dataframe prior to graphing, once we did so, we could plot directly without extracting the data as series.

Let's keep practicing, this time with histograms. 

9. Reproduce the histogram from Question 3 with the updated title and labels. This time change the color  of the bars to `"green"`. 
Use the dataframe prepared for you below.

```{python beach-max-wind-hist-pandas-setup, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="packages"}
no_outlier_wind = beach[beach['Maximum Wind Speed'] < 100][['Maximum Wind Speed']] # dealing with weird outliers
```
```{python beach-max-wind-hist-pandas, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-max-wind-hist-pandas-setup"}
...
plt.title("Distribution of Max Winds")
plt.xlabel("Max Wind in MPH")
plt.ylabel("Count")
```
```{python beach-max-wind-hist-pandas-solution, message = FALSE, warning = FALSE, echo = FALSE}
no_outlier_wind.plot.hist(color="green", bins=10)
plt.title("Distribution of Max Winds")
plt.xlabel("Max Wind in MPH")
plt.ylabel("Count")
```

```{r beach-max-wind-hist-pandas-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Histogram :)"
)
```

Notice how we didn't even have to pass any data to `.hist`! As long as you want to plot every column in the dataframe, calling `df.hist` is sufficient.
Let's use this to our advantage to make a overlaid histogram.

10. Make an overlaid histogram of both the maximum and hourly wind distributions. Let Matplotlib generate the colors automatically and use 10 bins. 
Use the updated `no_outlier_wind` data frame created below.


```{python beach-overlaid-wind-hist-pandas-setup, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="packages"}
no_outlier_wind = beach[beach['Maximum Wind Speed'] < 100][['Wind Speed', 'Maximum Wind Speed']] # dealing with weird outliers
```
```{python beach-overlaid-wind-hist-pandas, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-overlaid-wind-hist-pandas-setup"}
...
plt.title("Distribution of Max Winds")
plt.xlabel("Max Wind in MPH")
plt.ylabel("Count")
```
```{python beach-overlaid-wind-hist-pandas-solution, message = FALSE, warning = FALSE, echo = FALSE}
no_outlier_wind.plot.hist(bins=10)
plt.title("Distribution of Max Winds")
plt.xlabel("Max Wind in MPH")
plt.ylabel("Count")
```

```{r beach-overlaid-wind-hist-pandas-code-check, message = FALSE, warning = FALSE}
grade_this_code(
  correct = "Histogram :)"
)
```

While it's easy to overlay histograms and bar charts in `pandas`, it can be more challenging to do so with line plots. That's where packages like `seaborn` which further build on `matplotlib` and `pandas` can be helpful. Check out the [seaborn documentation](https://seaborn.pydata.org/generated/seaborn.lineplot.html), then try the following question.

11. Recreate the overlaid line plot from question 7 using seaborn. Let the legend go where ever seaborn thinks is best.

```{python beach-line-overlaid-seaborn-setup, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="packages"}
beach_day_max = beach[beach['Station Name'] != '63rd Street'][['Day of Year', 'Air Temperature', 'Station Name']].groupby(['Day of Year', 'Station Name']).max().reset_index().sort_values('Day of Year')
```

```{python beach-line-overlaid-seaborn, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="beach-line-overlaid-seaborn-setup"}

```
```{python beach-line-overlaid-seaborn-solution, message = FALSE, warning = FALSE, echo = FALSE}
sns.lineplot(beach_day_max, x="Day of Year", y="Air Temperature", hue="Station Name")
```
```{r beach-line-overlaid-seaborn-code-check, message = TRUE, warning = FALSE}
grade_this_code(
  correct = "Great job with Seaborn!",
)
```

For one last challenge with seaborn, we'll take a look at scatterplots. Checkout the [relevant documentation](https://seaborn.pydata.org/generated/scatterplot.html) if you want a hint.

12. Create a scatterplot to visualize the relationship between wind speed and air temperature, using differently shaped points for each beach.

```{python beach-scatter-overlaid-seaborn, include=TRUE, exercise = TRUE, message = FALSE, exercise.setup="packages"}
sns._____(data=beach, ______)
```
```{python beach-scatter-overlaid-seaborn-solution, message = FALSE, warning = FALSE, echo = FALSE}
sns.scatterplot(beach, x="Air Temperature", y="Wind Speed", style="Station Name")
```
```{python beach-scatter-overlaid-seaborn-hint, message = FALSE, warning = FALSE, echo = FALSE}
checkout the `style` argument of sns.scatterplot
```

```{r beach-scatter-overlaid-seaborn-code-check, message = TRUE, warning = FALSE}
grade_this_code(
  correct = "Great job with Seaborn!",
)
```

As we can see from this plot, there is a clear error with the readings at Foster Beach last year. 
Plotting these sorts of graphs early in the data exploration process can make it simpler to identify data quality issues before conducting more analysis.


## Grammar of Graphics Approach

So far we have used `Matplotlib`, `Pandas`, and `Seaborn`, which have afforded us varying degrees of control over the image being created. 
Now we are going to shift towards utilizing the grammar of graphics and packages like `altair` and `plotnine` to create graphs. 

Below we'll reproduce some of our graphs using the grammar of graphics. The key idea behind this approach is that every plot can be broken down into the **dataset**, the **mapping** between features of the dataset and visuals, and the **geometry** used to encode the observations. 
The grammar of graphics also has some nuance surrounding the coordinates and scales of the plot that we don't have time to address here. If you're interested, check out Leland Wilkinson's book *The Grammar of Graphics*.

For the remainder of this section, simply run the code. As you do so, compare it to the earlier approaches and focus on how we're able to do the same tasks with a different library.

### Building Up a Bar Chart

First we need to map the x and y axis to their relevant data.
```{python beach-bar-p9-aes, exercise=TRUE, exercise.eval = FALSE, message = FALSE, exercise.setup="packages"}
p9.ggplot(beach, p9.aes(x = 'Station Name', y='Wind Speed')).show()
```
Once we've done that we can use the `stat_summary` method to do all the computation we want for the bar chart. This makes the process much smoother than doing the computations separately. Note we assign `geom`="bar" for bar charts.

```{python beach-bar-p9, exercise=TRUE, exercise.eval = FALSE, message = FALSE, exercise.setup="packages"}
plot = p9.ggplot(beach, p9.aes(x='Station Name', y='Wind Speed')) + \
    p9.stat_summary(fun_y=np.mean, geom='bar', fill='skyblue', color='black') + \
    p9.labs(title="Average Wind by Beach")

plot
```
In the above plot we used `labs` to add a title, and the `fill` and `color` arguments to adjust the appearance of the graph. 
Feel free to experiment with those arguments to get the graph looking how you prefer. 

If you're really into colors check out this [list of color options](https://matplotlib.org/stable/gallery/color/named_colors.html) from Matplotlib!