---
title: "Introduction to BASH"
output: 
   learnr::tutorial:
      css: css/custom-styles.css
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
#knitr::knit_engines$set(python = reticulate::eng_python)

## Directory Requirements

#setwd("~/sandbox")

## Package Requirements for Lab

library(gradethis)
library(learnr)
library(tidyverse)

## Package Requirements  for Content

## Requirements for Answer Checking

custom_checker <- function(label, user_code, solution_code, envir_result, evaluate_result, last_value, stage, ...) {
  if (stage == "code_check") {
      if (user_code == solution_code){
          return(list(message = random_praise(), correct = TRUE))
      }
    return(list(message = random_encouragement(), correct = FALSE))
  }
}

gradethis_setup()
tutorial_options(exercise.completion = FALSE, exercise.checker = custom_checker)
```

```{bash setupbash}
cd data
```

```{r header, echo = FALSE}
library(htmltools)

tags$div(
  class = "topContainer",
  tags$div(
    class = "logoAndTitle",
    tags$img(
      src = "./images/dsi_logo.png",
      alt = "DSI Logo",
      class = "topLogo"
    ),
    tags$h1("Bash", class = "pageTitle")
  )
)
```


## Goals

The goal of this lab is to practice basic BASH commands using the command line interface. Please note that this lab borrows HEAVILY from the [Command Line Murder Mystery](https://github.com/veltman/clmystery) by Noah Veltman--once you have worked through the basic commands, you can either solve the murder mystery on the Shiny Server OR download the file and practice using the command line interface on your own machine!

## Setup

## Command Line Interface and BASH

### What is the command line interface?

The command line (also known as the command line interface, or CLI, or sometimes the terminal), is a **plain text-based interface** for executing commands on a computer. If you've ever seen a movie about hackers from the 1980s, like WarGames, where they stare at a prompt on a black screen and type in commands one at a time, it's basically that.

You have a prompt, and you can type in a command and hit 'Enter' to execute it. An example command would be:

```{bash example1, echo = TRUE, eval = FALSE}
touch newfile.txt
```

This command will create a file called `newfile.txt`.

### How do we access the command line?

You can open the command line differently according to different operating systems:

#### Mac OS X

Go to "/Applications/Utilities" and click on "Terminal" or search for "Terminal" in Spotlight (the magnifying glass at the top right corner of your screen).

#### Windows

Windows is a bit of a special case. If you go to the Start Menu and click "Run", and then type "cmd" and hit enter, it will open the Windows version of the command line. Unfortunately, the Windows version of the command line kind of has its own
system, so for the purposes of following these examples, you'll want to install [Cygwin](http://www.cygwin.com/), which will allow you to mimic a Linux-style
command line.

#### Desktop Linux

You can search for the "Terminal" application from the Dash. Let's be honest, though, if you're running Linux, you probably don't need this tutorial.

### BASH Syntax

BASH is one of the most commonly used languages used in the command line interface. BASH commands generally take the format:

```{bash example2, echo = TRUE, eval = FALSE}
[name of the command] [option] [option] [option] ...
```

Note that this is similar to specifying arguments inside a Python function or method.

### Directories and Paths

The prompt will also show what directory (folder) you're currently sitting in. Whenever you execute a command, you do it from a particular directory. This matters because when you execute a command that involves a filename or a directory name, you can
specify it one of two ways.

#### Relative Paths

Specifying a file or directory as a relative path means you are specifying where it sits *relative to the directory you're in* . For example, let's say you're in the `videos` subdirectory of the `files` directory. You'll see this prompt:

```{bash example3, echo = TRUE, eval = FALSE}
/files/videos$
```

Any commands you want to use need to go after the dollar sign, `$`. If you execute a command like `touch newfile.txt`, it will create `newfile.txt` inside the current directory. Relative paths don't start with a slash.

#### Absolute Paths

Specifying a file or directory as an absolute path means you are specifying where it sits *on the computer in absolute terms*, starting from the top level. For example, let's say you're in the `videos` subdirectory of the `files` directory again. 

```{bash example4, echo = TRUE, eval = FALSE}
/files/videos$
```

If you execute a command like `touch /files/music/newfile.txt`, it will create `newfile.txt` inside a different folder, the `music` subfolder of the `files` folder. Absolute paths start with a slash.

If you use an absolute path, the command will do the same thing no matter what directory you execute it from. So these two commands will have the same result from the `/files/videos` directory:

```{bash example5, echo = TRUE, eval = FALSE}
/files/videos$ rm video.mp4
```

This will delete the file `video.mp4` from the current directory.

```{bash example5_5, echo = TRUE, eval = FALSE}
/files/videos$ rm /files/videos/video.mp4
```

This will delete `video.mp4` from the `/files/videos/` directory, which happens to be the current directory.

The same two commands will not have the same result if you are in a different directory:

```{bash example6, echo = TRUE, eval = FALSE}
/files/text$ rm video.mp4
```

This will try to delete the file `video.mp4` from the `text` subdirectory instead, because that's the current directory.

```{bash example6_5, echo = TRUE, eval = FALSE}
/files/text$ rm /files/videos/video.mp4
```

This will delete the file from the `/files/videos/ directory`, even though it isn't the current directory).

**Remember: Starting a path with a slash means you want to give the entire path and ignore what directory you're currently in. Not starting a path with a slash means you want to give the path starting from the directory you're in.**

## Basic BASH Commands

You have already seen `touch`, but let's learn some other useful commands. 

### Print Working Directory

If you're ever unsure of what directory you're in, you can use the `pwd` (`p`rint `w`orking `d`irectory) command to get the absolute path of the current directory.

```{bash example7, echo = TRUE, eval = FALSE}
pwd
```

In the cell below, print your current working directory. 

```{bash ex1, exercise = TRUE, eval = FALSE}

```

```{bash ex1-solution, message = FALSE, warning = FALSE, echo = FALSE}
pwd
```

```{r ex1-code-check, message = FALSE, warning = FALSE}
grade_this_code()
```

Because we are on the server, you have a temporary directory with a complicated name--but you can navigate just as if you were working on your computer with permanent directories.

### Deleting Files

If you want to delete a file, you can use the `rm` (I think of it as `r`e`m`ove) command. Note that this file works in conjunction with relative and absolute paths. 
Say we have a specific file in our working directory named `specific_file.txt`. Removing this specific file looks like 

```{bash example7, echo = TRUE, eval = FALSE}
rm specific_file.txt
```

### File Patterns

In most cases when you have to specify a file name or directory name, you can also specify a general pattern that might match multiple files. There are lots of ins and outs with this, but the most basic version is using the asterisk (`*`), which matches anything. It's also known as a wildcard.

To delete any file in your current directory, run

```{bash example8, echo = TRUE, eval = FALSE}
rm *
```

To delete any text file (a file ending with the `.txt` extension), run

```{bash example9, echo = TRUE, eval = FALSE}
rm *.txt
```

To delete any file with a name starting with `data`, run

```{bash example9, echo = TRUE, eval = FALSE}
rm data*
```

Obviously, you can change the phrase or extension you are looking for! However, note that deleting files using the wildcard may remove something you actually wanted to keep, so be careful!

```{r q1, echo=FALSE}
question(
  "How would you delete all PDFs in your working directory?",
  answer("`rm *`"),
  answer("`rm *.txt`"),
  answer("`rm *.pdf`", correct = TRUE),
  answer("`rm pdf*`"),
  answer("Delete one specific `.pdf` file at a time until everything is gone.")
  allow_retry = TRUE
)
```

### Navigating

The two core commands for navigating what directory the prompt is in are `cd` and `ls`.

#### Changing Directories

`cd` is a command to change the current directory, and must be followed by a directory you want to change to. You can supply an absolute or relative path.

For example, to move from `files` to `files/videos`, run

```{bash example10, echo = TRUE, eval = FALSE}
cd videos
```

Note that when you change a directory on your computer, you will notice that the directory listed before the dollar sign, `$`, will change!

Once you change into a subdirectory, you can keep going. For example, to move to `videos` and then to `vines`, 

```{bash example11, echo = TRUE, eval = FALSE}
cd /videos
cd vines
```

You can jump multiple levels at once if you want.

```{bash example12, echo = TRUE, eval = FALSE}
cd /videos/short
```

Make sure to type the name of the directory exactly, otherwise you will get a message like 

```{bash example13, echo = TRUE, eval = FALSE}
cd: no such file or directory:
```

This can also happen when you supply the absolute file name but forget a subdirectory. There are a couple of habits you should establish to make sure this doesn't happen:

1. Be careful when typing (obvious, but harder than you think when you get very complicated names!!) 
2. Copy and paste the path/filename--on a Mac, right click on the file or directory you are interested in using Finder and then click on "Get Info". On a PC, right click on the file or directory and then click on "Properties."
3. Use the "Tab" key for autocompletion. This will feel weird at first but can dramatically speed up your work!
4. Use the up arrow key to find previous commands--if you visit a directory a lot, you'll probably find it again!

You can also move up a directory, using

```{bash example14, echo = TRUE, eval = FALSE}
cd ..
```

#### Listing Files

`ls` will `l`i`s`t the files in the current directory. It's helpful for figuring out where you are, what files exist, and what subfolders exist.

```{bash example15, echo = TRUE, eval = FALSE}
ls
```

Using `ls -l` will print the list vertically, with lots of other extra information about the file size, permissions, and last modified date--the command line interface isn't always very readable, so it's helpful to use tips like this when you can. 

### `head`

The head command will print out the first 10 lines of a file:
/files$ head names.txt
You can also specify a different number of lines. This will print out the first 15 lines of a file:
/files$ head -n 15 names.txt
Or, if you want to print all the file but leave out the LAST 15 lines, you can give a negative number:
/files$ head -n -15 names.txt
One of the nice uses of head is to quickly peek inside a large text file to see what's in it without having to wait for a text editor to
load it. This becomes a big deal when you're talking about a 1 GB file!

### `tail`

The tail command is the reverse of head. It will print out the last 10 lines of a file:
/files$ tail names.txt
This will print out the last 15 lines of a file:
/files$ tail -n 15 names.txt
Or, if you want to print all the file but leave out the FIRST 15 lines, you can add a plus sign:
/files$ tail -n +16 names.txt
This is helpful if you want to, say, remove a header row from a CSV file:
/files$ tail -n +2 names.txt > names-no-header.txt

### grep

The grep your search.
command will let you search a file (or multiple files) for a phrase. By default, it will print out each line that matches
Print out lines that contain the word "darkwing":
/files$ grep "darkwing" famousducks.txt

Same as above, but the search is case-insensitive:
/files$ grep -i "darkwing" famousducks.txt
Find matches for the exact word "Donald" in a file - words that contain "Donald," like "McDonald," won't count:
grep -w "Donald" famousducks.txt
Find matches for "McDuck" in every file in the current directory:
grep "McDuck" *
Find matches for "McDuck" in every file in the current directory AND every subdirectory, all the way down:
grep -r "McDuck" *
For each match of "Howard", print out that line AND the 4 lines after it (5 lines total):
grep -A 4 "Howard" famousducks.txt
For each match of "Howard", print out that line AND the 4 lines before it (5 lines total):
grep -B 4 "Howard" famousducks.txt
For each match of "Howard", print out that line AND the 4 lines before it AND the 4 lines after it (9 lines total):
grep -C 4 "Howard" famousducks.txt
Instead of printing out the matching lines themselves, print out the filenames that match your search:
grep -l "Daffy" *
Just get the number of matches:
grep -c "Daffy" *
Show line numbers along with the matching lines:
grep -n "Daffy" famousducks.txt

### pipe

You can send the output to another command using the pipe symbol (|). The grep command searches through some text for
matches (more on this later), so you could do this to get the first 10 lines of a file, and then search for "Steve" within those 10
lines:
/files$ head names.txt | grep "Steve"
This is basically the same as doing this:
/files$ head names.txt > temporaryfile.txt
/files$ grep "Steve" temporaryfile.txt
But instead of first sending the output to a file and then running the second command on that file, you pipe the output directly
from the first command into the second. You can chain as many of these together as you want:
/files$ grep "United States" addresses.csv | grep "California" | head
This would search the file addresses.csv for lines that contain the phrase "United States", then search the results for lines that
contain the word "California", and then print out the first 10 of those matches.

### cat

The cat command will combine multiple files together. This will print three files in a row, as if they were one file:
cat turkey.txt duck.txt chicken.txt
Remember that this will just print the output into your terminal.

### Regular Expressions

When using something like grep to search, you can search for a simple term with only letters, numbers, and spaces. But if you
want to search for a pattern, you can use what's called a regular expression. Regular expressions use special characters to
represent patterns, like "any number," "any letter," "X or Y," "at least three lowercase letters," and so on.
We won't worry about the ins and outs for now, but one useful operator is the period (.). In regular expression-ese, this means
"One of any character." So you can search for something like:
/files$ grep -i "car.s" dictionary.txt
This would match words like cards , carts , cares , and so on. It would also match the middle of the phrase "scar story" (CAR
S) because "any character" means ANY character, including a space or a punctuation mark.
One more example:
/files$ grep -i ".e.st" dictionary.txt
This would match things like least , beast , and heist .

### Command List

So, for reference, here are all the commands we've covered:

* `touch`: Creates a new file
* `pwd`: Prints your working directory. 
* `cd`: Changes your working directory. 
* `ls`: Lists the files in your working directory. 

## Murder Mystery

## Python Specific Tasks

* installing packages
* running scripts

## Resources
