---
title: "Lists, Arrays, and Dictionaries Lab"
output: 
   learnr::tutorial:
      css: css/custom-styles.css
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

library(shiny)
library(gradethis)
library(learnr)
library(reticulate)

custom_checker <- function(label, user_code, solution_code, envir_result, evaluate_result, last_value, stage, ...) {
  if (stage == "code_check") {
      if (user_code == solution_code){
          return(list(message = random_praise(), correct = TRUE))
      }
    return(list(message = random_encouragement(), correct = FALSE))
  }
}

gradethis_setup()
tutorial_options(exercise.completion = FALSE, exercise.checker = custom_checker)
```

```{r header, echo = FALSE}
library(htmltools)

tags$div(
  class = "topContainer",
  tags$div(
    class = "logoAndTitle",
    tags$img(
      src = "./images/dsi_logo.png",
      alt = "DSI Logo",
      class = "topLogo"
    ),
    tags$h1("Data 119 - Python Data Storage Review", class = "pageTitle")
  )
)
```

# Goals

The goal of this lab is review some of what you've learned about coding in Python. We will cover:

- Lists
- Arrays
- Dictionaries
- Sets
- Tuples

# Storing Data in Python

Python and its libraries provide many ways to group data together.   Some important ones:

- Lists, Tuples, Sets, Dictionaries (built-in to Python)

- Arrays (found in the `NumPy` library)

- DataFrames (in `Pandas`)


The above are listed in order of increasing functionality and sophistication

In general you should use the simplest one that meets your needs...


## Sequences (built into Python)


Sequences are a basic type of structure to group data. You don't need to import any additional libraries to use them.

We will focus mostly on one type of sequence: **lists**! (**sets** and **tuples** are other useful types of sequences)

A **list** is an ordered sequence of values that can be changed.

You can create a list by using square brackets "[]". Each element is separated by commas.

### Let's create a list!

```{python list1, exercise=TRUE, exercise.eval = FALSE}
School_locations = ["California", "Illinois", "North Carolina", "Texas", "Georgia", "Washington D.C."]
School_locations
```

#### Lists can have different data types!

```{python list2, exercise=TRUE, exercise.eval = FALSE}
Random_stuff = ["California", 38, 3.14159] #A string, int, float
Random_stuff
```

### Built-in Functions

Recall Python has built-in functions of the form `function_name()`
A very common one is the `print()` function!

`type()` is another important function. We can call this on a number, string, list, or any object to see type it is!

```{python listfuncs, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'list2'}
print("California")
print(type("California"))
print(type(38))
print(type("38"))
print(type(3.14))
print(type(Random_stuff))
```

We can also find the length of a list:

```{python listfuncs2, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'list2'}
len(Random_stuff)
```


## Accessing values from a list


Each element of a sequence is assigned an index corresponding to its position where **indices start at 0**. We can access an element by calling the sequence or list and putting in square brackets the element we want!

```{python listvals1, exercise=TRUE, exercise.eval = FALSE}
School_locations = ["California", "Illinois", "North Carolina", "Texas", "Georgia", "Washington D.C."]
School_locations[1]
```

Negative indexes count from the end of the sequence

```{python listvals2, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'listvals1'}
School_locations[-1]
```

If our list is long or we want the first or second from end this is useful!


1. What will be the result of running the following code? Answer in the box below.

```python
School_locations = ["California", "Illinois", "North Carolina", "Texas", "Georgia", "Washington D.C."]

School_locations[6]
```

```{r q1, echo=FALSE}
question(
  "What will be the result of running the code above?",
  answer("California"),
  answer("Illinois"),
  answer("North Carolina"),
  answer("Texas"),
  answer("Georgia"),
  answer("Washinton D.C."),
  answer("An Error", correct = TRUE),
  allow_retry = TRUE
)
```


### List Slicing

We can also extract a "slice" of a list.

The range of elements can be specified with colons. The output is a list starting at left index and stopping at (right index -1). The range you specify is a half-closed interval \[start,end)

```python
list[start: end]
```

```{python listslice1, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'listvals1'}
School_locations[1:3] # up to but not including the end of the slice
```

The above prints out the second and third elements.

We can still slice a list by 'leaving out' a starting (or ending) position. The missing position will revert to the default. The default start value is the beginning of the list. The default end value is the end of the list.

```{python listslice2, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'listvals1'}
School_locations[:3] #start at beginning stop at index end-1 -----> 3-1=2
```

2. What will the following output? Answer in the box below.

```python
School_locations = ["California", "Illinois", "North Carolina", "Texas", "Georgia", "Washington D.C."]  

School_locations[:]
```

```{r q2, echo=FALSE}
question(
  "What will be the result of running the code above?",
  answer("[]"),
  answer("[California, Washington D.C.]"),
  answer("[California, Illinois, North Carolina, Texas, Georgia, Washington D.C.]", correct = TRUE),
  answer("An Error"),
  allow_retry = TRUE
)
```


### Default Settings: List slicing

It might be useful to take out every even index from a list.

There is an optional argument we can use when slicing a list, the step.
In general we can slice by using ```list[start: end: step]``` where the default (what is used when it isn't specified) of step is 1

```{python listslice3, exercise=TRUE, exercise.eval = FALSE}
School_locations = ["California", "Illinois", "North Carolina", "Texas", "Georgia", "Washington D.C."]

School_locations[::2]
```

3. What do you think the following will output? Answer in the box below.

```python
School_locations[:3:3]
```

```{r q3, echo=FALSE}
question(
  "What will be the result of running the code above?",
  answer("[]"),
  answer("[California]", correct = TRUE),
  answer("[California, Texas]"),
  answer("[North Carolina, Washington D.C.]"),
  answer("[North Carolina]"),
  answer("[California, Illinois, North Carolina, Texas, Georgia, Washington D.C.]"),
  answer("An Error"),
  allow_retry = TRUE
)
```

### Operations and manipulation on lists

We can: 
- Insert items into lists!
    - at either end
    - in the middle
         
- Count how many items have a specific value

- Sort elements

How do we do this? By using methods. 

*Methods* are particular built-in functions that work on objects in python. There are specific methods that work for all *list* objects!

Methods take the form:

```python
list.method()
```

Built-in functions can also be applied to objects in python. Recall they take the form:

```python
function_name(list)
```

### Appending to a list

You can append an item to the end of a list:

```{python append1, exercise=TRUE, exercise.eval = FALSE}
School_locations = ["California", "Illinois", "North Carolina", "Texas", "Georgia", "Washington D.C."]

School_locations.append('Michigan')
School_locations
```

Or you can insert a value at a particular index:

```{python append2, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'append1'}
School_locations.insert(4,'tomato')
School_locations
```

Note we didn't assign the result to `School_locations`, but this automatically changes/overwrites the object.

Let's make a copy of our list so we can see how this works.

```{python append3, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'append1'}
School_location_og = School_locations.copy()
School_location_og
```

4. Append the number 2023 to our list and then view the list. Note even though the list contains only strings, I can append an integer as lists accept any data type.

```{python append4, exercise=TRUE, exercise.eval = TRUE, exercise.setup = 'append2'}

```

```{python append4-solution, message = FALSE, warning = FALSE, echo = FALSE}
School_locations.append(2023)
School_locations
```

```{r append4-code-check, message = FALSE, warning = FALSE, echo = FALSE}
grade_this_code()
```

This does not change our copy of the list.

```{python append5, exercise=TRUE, exercise.eval = FALSE, exercise.setup = 'append3'}
School_location_og
```

### How do we find methods?

Use online [documentation](https://docs.python.org/3/tutorial/datastructures.html)!

Or use the built-in function ```dir()```.

```{python dir, exercise=TRUE, exercise.eval = FALSE}
print(dir(School_locations))

print(dir(list))
```

### Let's play with lists!

Create a list with at least 4 items. (Recall items can be a string, int, float.) Name the list `my_list`.

```{python ex1, exercise=TRUE, exercise.eval = FALSE}
#Define your list here
```

5. Remove elements 2 and 3

```{python ex2, exercise=TRUE, exercise.eval = TRUE, exercise.setup = 'ex1'}
#Code here
```

6. Append your favorite number to the end of the list

```{python ex3, exercise=TRUE, exercise.eval = TRUE, exercise.setup = 'ex2'}
#Code here
```

7. Insert the string "Math is fun" at index 2

```{python ex4, exercise=TRUE, exercise.eval = TRUE, exercise.setup = 'ex3'}
#Code here
```

8. 

```{r q3, echo=FALSE}
question(
  "Did your above operations change your original list, or create a copy?",
  answer("Change Original", correct = TRUE),
  answer("Create a Copy"),
  allow_retry = TRUE
)
```

9. Given list ```number_list = [1,2,3,4,5,6,7]```, create a new list ```new_list = [6,4,2]``` in 2 lines of code. You may have to use list methods.

```{python ex5, exercise=TRUE, exercise.eval = TRUE}
#Code here
```

## Grouping Data using Arrays

Arrays (from the `NumPy` library) are another way to collect data. Like a list, they contain a sequence of values. But, unlike a list, all elements of an array must have the same data type. This is because `NumPy` arrays were built for efficient computation. They can perform operations on all elements in one step. They can also do element-wise computation. This means that, for example, if you add two arrays together, the result will be an array where the element at each index is the sum of the elements at the index in the two original arrays. For this reason, when two arrays are added (or subtracted, multiplied, or divided) they must have the same size.

### Lists vs Arrays

Lists are more flexible
  - Can contain elements of different types
    
`NumPy` arrays have some advantages
  - size - they take up less computer memory than lists
  - performance - faster access than lists
  - functionality - linear algebra functions built-in
  - can be multiple dimensions

### Creating an array

First import the `NumPy` library!! We do this using an import statement and we give `NumPy` the alias `np`. This makes it so that any time we want to access the `NumPy` library, we only need to type `np`.

We can create an array using `np.array()` which takes in a list of values of the same data type. Below is an example of writing the import statement and creating an empty array. The array is empty because the list has no elements.

```python
import numpy as np

np.array([])
```

Now, let's actually import numpy and create an array with two elements.

```{python array1, exercise=TRUE, exercise.eval = FALSE}
import numpy as np

np.array([1,2])
```

Below, we create a list that contains integers and a string.

```{python array2, exercise=TRUE, exercise.eval = FALSE, exercise.setup = "array1"}
tomato_list = [22, 38, 26, 35, 35,'tomato']
print(tomato_list)
```

Let's see what happens when we try to make an array from this list.

```{python array3, exercise=TRUE, exercise.eval = FALSE, exercise.setup = "array2"}
tomato_array = np.array([22, 38, 26, 35, 35,'tomato'])
print(tomato_array)
```

A list can have different types, but an array will default to one. So you can see all `int`s were changed to `strings` (notice the single quotes around the numbers).

An array will make sure everything is the same type.

### Exploring arrays

Let's create an array by converting a list to a `NumPy` array.

```{python prime50, exercise=TRUE, exercise.eval = FALSE}
prime50_array = np.array([2, 3, 5, 7, 11, 13, 17, 19, 23, 29, 31, 37, 41, 43, 47])

print(type(prime50_array))
```

Extraction and slicing of one dimensional arrays work exactly the same as lists!

10. Print the second element of `prime50_array`.

```{python prime50, exercise=TRUE, exercise.eval = TRUE}
print(prime50_array[1])
```

11. Print the array starting at index 1 and up to (but excluding) index 2.

```{python prime50, exercise=TRUE, exercise.eval = TRUE}
print(prime50_array[1:2]) 
```


12 Will the following give the same or different outputs?

```python
prime50_array[1::2]

prime50_array[::2]
```

Answer here

### Arrays have attributes

**Attributes** are characteristics of an object. We can view an object's attributes by using the dot operator `.` similarly to when we used methods by using the syntax `object.method()`. However, to access an attribute, we don't use parentheses.
(See https://numpy.org/doc/stable/reference/generated/numpy.ndarray.html)

Two attributes of arrays are size, which we can access using `.size`, and shape, accessed using `.shape`.

```{python}
prime50_array.size
```

```{python}
prime50_array.shape
```

### Arrays have useful methods

Like lists, there are a lot of useful methods we can use on array objects. Here are a couple:

 -  `.sum()`
    
 -  `.mean()`

 -  `.nonzero()`


(see the documentation for a complete list: https://numpy.org/doc/stable/reference/generated/numpy.ndarray.html)

We can sum arrays, but we cannot always sum a list because a list can have different data types!

```{python}
print(prime50_array.sum())
```

## Numpy also has many useful built-in functions to use on arrays!

Notice there is often more than one way to do a common operation!

```{python}
print(prime50_array.sum()) #sum method

print(np.sum(prime50_array)) #sum function (from NumPy library)

print(np.count_nonzero(np.array([1,2,0,2,1,0,2])))

print(prime50_array.mean()) #mean method

print(np.mean(prime50_array)) #mean function
```

### Printing Formatted Code


We often find that we want to format the output of our code in a nice way. This includes printing string and code output. In Python, there are multiple ways to do this. Two options are listed below. Consider our prime50_array for this exercise.

```{python}
prime50_array
```

```{python}
#Option 1: print string and code seperated by a comma
print("array:", prime50_array)

print("max:", np.max(prime50_array)) #we can apply functions directly to code output


#Option 2: print string and code by converting everything to a string and concatenating
print("Maximum prime under 50 is " + str(np.max(prime50_array)))
```

### ✏ <mark style="background-color: Thistle"> Code Comprehension: </mark>

Take the last output and add a period in the code output.


### We can easily create arrays by specifying a range.

Calling ```np.arange()``` creates a half-closed interval \[start,end) - the end value is not included


```{python}
np.arange(4,10)
```

If you leave out the start, the default is zero.

```{python}
print(np.arange(10))
```

We can specify a step size we want to increment by. If we leave out the step, the default is one

```{python}
print(np.arange(1,31,2))
```

## Another reason why arrays are useful!

Elementwise operations!

```{python}
array_1 = np.arange(10)
array_2 = np.array([1,2,3,4,5,6,7,8,9,10])
difference_array = array_1 - array_2
difference_array
```

## ✏ <mark style="background-color: Thistle"> Code comprehension: Multiple Choice </mark>

#### What will be printed?

```python
import numpy as np
a = np.array([1,2,3,5,8])
b = np.array([0,3,4,2,1])
c = a + b
c = c*a
print (c[2])
```

A. 7

B. 12

C. 10

D. 21

E. 28

## ✏ <mark style="background-color: Thistle"> Code comprehension: Multiple Choice </mark>

What will be output for the following code?

```python
 number_array = np.array([1,2,3,5,8])
 number_array = number_array + 1
 print(number_array[1])
```

 A. 0
    
 B. 1

 C. 2
    
 D. 3

### We may have higher dimensional arrays.


Organized into rows and columns

```{python}
arr_2d = np.array([[1, 2, 3], [4, 5, 6]]) #specify first row then second

print(arr_2d)
```

```{python}
arr_2d.shape #outputs (#rows, #columns)
```


We can reshape our arrays

```{python}
reshaped_arr = np.reshape(arr_2d, 6)

reshaped_arr.shape
```

```{python}
np.reshape(arr_2d, (3,-1)) # the unspecified value is inferred to be 2

np.reshape(arr_2d, (3,-1)).shape
```

```{python}
arr_2d[:,2] #all of the rows, column at index 2
```

```{python}
arr_2d[0,1] #element with row index 0 and column index 1
```

## ✏ <mark style="background-color: Thistle">Working with Arrays: Exercise</mark>

Use the following array to answer the questions

```{python}
random_number_array = np.array([32, 56, 78, 3, 15, 109, 13, 24, 58, 61, 90, 93, 45, 21, 46])
```

1. Remove elements 2 and 3

```{python}
#code here
```

2. Use a method to find the minimum value in the array

```{python}
#code here
```

3. Find the 4th smallest element in the array

```{python}
#code here
```

4. Create an array with 60 elements that corresponds to the 60 minutes in an hour.

```{python}
#code here
```

5. Starting (and including) minute 4, extract every 5th element.

```{python}
#code here
```

## Another Built-in Collection of Data: Dictionaries

A **dictionary** is a set of "key: value" pairs where each key is unique.

We can create a dictionary with curly brackets "{}"

Entries of a dictionary are of the form "key: value"

```{python}
survey_dict = {0: "Strongly Disagree", 1: "Disagree", 2: "No opinion", 3: "Agree", 4: "Strongly Agree"}
```

We can access values of a dictionary by their key. (This is why all keys must be unique!)

```{python}
survey_dict[1]
```

Dictionaries are useful for storing and extracting data! Think of them like an address book where you look up someone's address by finding their name.

A few useful operations with dictionaries:

 - Add an entry

 - Delete an entry

Add and delete pairs!

```{python}
del survey_dict[1]
```

```{python}
survey_dict
```

```{python}
#this adds a key-value pair
survey_dict['new_key'] = 'new value'
```

```{python}
survey_dict
```

Note we can also determine if keys are contained in the dictionary

```{python}
3 in survey_dict
```

```{python}
"Disagree" in survey_dict
```

```{python}
'new_key' in survey_dict
```

Keep in mind, keys do not all need to be the same type..although it may make more sense to keep them that way.

```{python}
survey_dict_2 = {"Strongly Disagree": 0 , "Disagree": 1 , 2: "No opinion", 3: "Agree", 4: "Strongly Agree"}
```

```{python}
survey_dict_2["Disagree"]
```

```{python}
list(survey_dict_2)
```

### Dictionary Methods

There are methods that you can use on dictionaries, just like with lists and arrays. Here are a few.

```{python}
print(dir(survey_dict_2))
```

```{python}
survey_dict.keys()
```

```{python}
survey_dict
```

## ✏ <mark style="background-color: Thistle">Working with Dictionaries: Short Exercise</mark>

Below is a dictionary containing total number of homicides in the United States in 2021, by state. (Published by Statista Research Department, Oct 14, 2022). Note Washington D.C. is included as 'District of Columbia'.

1. Pick a few states of interest and find their homicide number (use code here...do not just manually search the dictionary!)


2. Find the number of keys in the dictionary. Does this imply all 50 states are included here?


3. What are some limitations to this data?

```{python}
homicide_dict = {'Texas': 2064, 'North Carolina': 928, 'Ohio': 824, 'Michigan': 747, 'Georgia': 728, 'Tennessee': 672, 'Missouri': 593, 'Virginia': 562, 'South Carolina': 548, 'Illinois': 514, 'Pennsylvania': 510, 'Louisiana': 447, 'Indiana': 438, 'Alabama': 370, 'Kentucky': 365, 'Colorado': 358, 'Washington': 325, 'Arkansas': 321, 'Wisconsin': 315, 'Oklahoma': 284, 'Nevada': 232, 'Minnesota': 203, 'Arizona': 190, 'Oregon': 188, 'New Mexico': 169, 'Mississippi': 149, 'Connecticut': 148, 'Maryland': 138, 'New Jersey': 137, 'Massachusetts': 132, 'New York': 124, 'California': 123, 'District of Columbia': 109, 'West Virginia': 95, 'Delaware': 94, 'Kansas': 87, 'Utah': 85, 'Iowa': 70, 'Rhode Island': 38, 'Idaho': 36, 'Montana': 31, 'South Dakota': 26, 'Nebraska': 25, 'Alaska': 18, 'Maine': 18, 'Wyoming': 17, 'New Hampshire': 14, 'North Dakota': 14, 'Vermont': 8, 'Hawaii': 6}
```

```{python}
#code here
```

## Grouping Data with DataFrames

One of the most commonly used collections of data in data science is the DataFrame. DataFrames come from the pandas library which we will talk in more detail about when you get to campus! For now, here is a quick introduction.

### Different objects for different goals

There are different collections of data in Python because each works well in different contexts. We've already talked about lists and arrays:

Method 1: (Using lists from python (no import numpy needed))
- list of lists
- Hard to manipulate


Method 2: (Using np.array)
- All values of same data type
- Easy to do math and matrix manipulations
- No row column names

DataFrames are useful if we have multiple observations of data of different types.

Imagine you are conducting a survey. On this survey you ask about a person's height in inches, favorite fruit, and whether they own a dog. Each of these pieces of information would be stored in Python using a different data type (✏ As an exercise, what data type would you use for each?)

Each person that you survey provides a different 'observation' of those 3 pieces of information. Picture a DataFrame as a table with rows and columns. Each row corresponds to a different person being surveyed and each column corresponds to their answer to a particular question. Each column contains the same data type within that column but individual columns can be of different data types. Here are some useful elements of DataFrames:

Method 3: (Using pandas DataFrames)
- Different columns can have different types
- Easy to manipulate by name
- row and column names built in!
- Rows are "individuals" or "instances"
- Columns are attributes of those individuals

### Storing Data in 3 ways

Suppose we take a survey of student ID, favorite number, and favorite food.

We can store this data as a list of lists, a 2-dimensional array, or a DataFrame.

```{python}
list_of_lists=[['Pizza','Pierogi','Ramen'],  #fav food
               [0,22,-3.1415],  #fav number
               [1234, 4456, 5882]]   #Student ID

list_of_lists
```

```{python}
#This is a 2D array
np.array([['Pizza','Pierogi','Ramen'],  #fav food
               [0,22,-3.1415],  #fav number
               [1234, 4456, 5882]]) #student id
```

```{python}
#ignore this code for now!
import pandas as pd
df = pd.DataFrame(
    { 1:['Pizza','Pierogi','Ramen'],  #fav food
      2:[0,22,-3.1415], #fav number
     3:[1234, 4456, 5882]}) #id

df
```

## Wrapping Up

In this notebook, we learned about lists, arrays, and dictionaries and were introduced to the concept of DataFrames. During the next lab, we will build on these concepts so that you feel comfortable working with and manipulating large groups of data. See you soon!

















## Conditional Statements

There are some situations where we want to proceed with a task or perform an action dependent on whether a certain condition, or perhaps conditions, have been satisfied.

Conditional statements have the form of an "if-then" statement *if* statement `P`, the hypothesis, occurs, *then* statement `Q`, the conclusion, also occurs.

How do we write this in code? We utilize the `if` expression in Python.

The statement below will execute the indented block *conclusion*, if the *hypothesis* is true; otherwise, if *hypothesis* is not true, then the indented block is ignored:

```python
if hypothesis:
    conclusion
```

We implement an example by checking if a given number is positive. The below code says, if the provided x is greater than 0, the indented body will run, and 'Positive' is printed. Otherwise (if x < 0) the indented body is not executed.

```{python}
if x > 0:
    print('Positive')
```

We can define a value for x and see what happens.

```{python}
x = 5

if x > 0:
    print('Positive')
```

```{python}
x = -6

if x > 0:
    print('Positive')
```

If we want to output something when we enter a negative number we can do the following:

```{python}
#we can keep our x value defined to be -6
x = -6

if x > 0:
    print('Positive')

if x < 0:
    print( 'Negative')
```

In the above code block, the first if condition is tested, it evaulated to False, and then the next if condition is tested, which evaulates to True, hence the body is executed - output is printed.

### We can use `elif` statement instead.

Used in combination with an `if` expression, an `elif` statement is only checked if all previous statements evaluate to False. This allows us to check if our first condition is true, otherwise we move to evaluate the truth value of the next statement.

```{python}
if x > 0:
    print('Positive')

elif x < 0:
    print('Negative')
```

We haven't addressed all possibilities! What happens when we define `x = 0`? Let's redefine the options to have an input in all cases.

```{python}
#Now we can define x = 0
x=0

if x > 0:
    print('Positive')

elif x < 0:
    print('Negative')

elif x == 0:
    print('Neither positive nor negative')
```

## Else statement

In an `elif` statement each condition is checked until a condition is true. Often we can replace the last `elif` statement with an `else` statement, whose body will be executed only if all the previous comparisons are false.

At this point, all other conditions have been evaluated and none executed. Thus, there is no condition or hypothesis associated with the `else` statement. If it is reached, the conclusion of the `else` statement is executed.

```{python}
if x > 0:
    print('Positive')

elif x < 0:
    print('Negative')

else:
    print('Neither positive nor negative')
```

## The General Format


```python
if hypothesis_1:
    conclusion_1
else:
    conclusion_2
```
    
Or we could have $n+1$ conclusions for a chosen $n$ in which case we have the format:
    
```python
if hypothesis_1:
    conclusion_1
elif hypothesis_2:
    conclusion_2
...
elif hypothesis_n:
    conclusion_n
else:
    conclusion    
```    

BE CAREFUL!!!  Since the `else` statement is executed without checking a condition, you want to be absolutely certain that all desired possibilies are accounted for in the previous condition(s).

### Exercise 1.
Use `if` statements to print your letter grade given a numerical grade where the following is true:

    (a) Above a 90 is an A
    
    (b) Between 80 and 90 is a B, including 80
    
    (c) Between 70 and 80 is a C, including 70
    
    (d) Between 60 and 70 is a D, including 60
    
    (e) Below a 60 is an F
    
Check your code on the following grades:

    my_grade = 83
    
    my_grade = 70
    
    my_grade = 57

```{python}
#code here
```

### Exercise 2
Describe the difference in behavior between the following pieces of code. HINT: Try giving different values of x and executing the cell.

```{python}
# CODE SEGMENT A
x = ...
if x > 0:
    print("x is positive")
elif x < -5:
    print("x is less than -5")
else:
    print("x is between -5 and 0")
```

```{python}
# CODE SEGMENT B
x = ...
if x > 0:
    print("x is positive")
if x < -5:
    print("x is less than -5")
else:
    print("x is between -5 and 0")
```

Your answer here:

## For Statements

We might be interested in repeating something 100, 1000 times, or performing some action for each element in a list or array. If this is the case, we want to use `for` statements.

A `for` statement can *iterate* through a sequence to perform some action on each of its elements. The general form of a `for` statement is below:

```python
for item in sequence:
    action    
```

And for each of the *items* in *sequence* the indented body of the `for` statement is executed (here "action").

 – or the indented body is "looped" –

This sequence can really be any *iterable* object, including a list, a string, or a range of numbers, to name a few.


Notice we specify a name to assign to the value of each of the sequence's items – here `item`. This name is assigned to each of the values `in` the sequence, sequentially.

## An example

```{python}
list_of_things= ["red", 2, 7.3, "dog"]
for element in list_of_things:
  print(element)
```

This is equivalent to the following code.

```{python}
element = "red"
print(element)

element = 2
print(element)

element = 7.3
print(element)

element = "dog"
print(element)
```

In fact for relatively simple for loops,

```python
for item in sequence:
    action    
```

is equivalent to

```python
item = sequence[0]
action    

item = sequence[1]
action

item = sequence[2]
action
```

etc.

### What can we iterate over?

Note that what we iterate over does not need to be directly related to the body of the `for` statement. In fact, `for` statements are useful to simply execute the body or action a given number of times.

In the first example below, the value of the *iterator* is used in the body of the `for` statement; in the second example, the `for` statement uses this iterator merely to repeat or loop through the body statement a certain number of times.

```{python}
for i in [1,2,3]:
    print(i)
```

We could also specify a *range* of values

```{python}
for i in range(5):
    print(i)
```

```{python}
for i in range(2,5): #we could choose the starting and ending values here too.
    print(i)
```

The iterator, i, does not have to appear in the body of the for statement, or for loop.

```{python}
for i in range(3):
    print('hello')
```

Often for loops are useful just to repeat something a specified number of times.

### Exercise 3
Convert the following piece of a code to a for loop that does the same thing.

```{python}
x = "I"
if len(x) > 3:
    print(x)

x = "only"
if len(x) > 3:
    print(x)

x = "use"
if len(x) > 3:
    print(x)

x = "long"
if len(x) > 3:
    print(x)

x = "words"
if len(x) > 3:
    print(x)
```

```{python}
# Write your code in this cell.
l = ["I","only","use","long","words"]
...
```

## A Common Pattern

A common type of task you may find yourself doing is using a for loop to build a list.

```python
result = []
for i in list:
  ...
  result.append(x)
```

The code below creates a list storing the first 100 perfect squares.

```{python}
squares = []
for i in np.arange(0,101):
  squares.append(i**2)
squares[0:10]
```

Similarly, the code below builds a list of the first 1000 fibonacci numbers.

```{python}

fibonacci = [0,1]
for i in np.arange(2,1001):
  fibonacci.append(fibonacci[i-1]+fibonacci[i-2])
```

## Warning

You may find yourself tempted to always use ``np.array`` for every list of numbers. However, they do have a disadvantage in this situation. It is significantly slower to append an element to a ``np.array`` than a list.

```{python}
from time import process_time
t0 = process_time()
test_list = []
for i in np.arange(0,100000):
  test_list.append(i)

print("It took ", process_time()-t0, " seconds.")
```

```{python}
t0 = process_time()
test_list = np.array([])
for i in np.arange(0,100000):
  test_list = np.append(test_list,i)

print("It took ", process_time()-t0, " seconds.")
```

## Nested *for* loops

Suppose we have two lists, and we want to pair every element in `list_1` with every element in `list_2`.

This takes the following form:

```python
for item_1 in list_1:
    for item_2 in list_2:
        print(item_1, item_2)
```

We *could* write out by hand all the possible combinations pairing elements of `list_1` with `list_2` … or, we could use nested `for` statements to systematically consider each element in `list_2` for each element in `list_1`.

### Example

Suppose we want to find all possible combinations of the below lists.

```{python}
my_animals = ['dog', 'cat', 'cow']

adjectives = ['hairy', 'scary', 'cute']

for adj in adjectives:
    for animal in my_animals:
        print(adj, animal)
```

Order of the inner and outer *for* loops does matter!

```{python}
for animal in my_animals:
    for adj in adjectives:
        print(adj, animal)
```

### Exercise 4

Write code to print the first 10 natural numbers (positive whole numbers starting at 1)

**Talk to your neighbor about your thought process**

```{python}
#code here
```

## Exercise 5

Use a for loop and the function ``np.sqrt`` to build a list containing the square roots of the first 10000 natural numbers.


### Exercise 6
Write code to find the sum of the numbers 1-100 (You should get 5050).


**Talk to your neighbor about your thought process**

```{python}
#code here
```

### Exercise 7
Write code to display the pattern below:

```python    
*
**
***
****
```


**Talk to your neighbor about your thought process**

```{python}
#code here
```

### Exercise 8
Write code to display the pattern below:

```python    
1
12
123
1234
```

**Talk to your neighbor about your thought process**

```{python}
#code here
```

### Exercise 9

Write code that will take a string, split it up into individual words, and count the number of capitalized words. It should work for any sentence s. HINT: Search the internet to find a string method that can determine if a world is capitalized or not.

```{python}
s = "This sentence written by Bill Trok in Chicago, Illinois has five capitalized words."
words = s.split()
# Your Code Here
```

