---
title: "Introductory Data Science Labs"
author: "Amy Nussbaum, Amanda Jotte, Atticus Ginsborg"
date: "2026-05-01"
format: 
  beamer:
    theme: "Berkeley"
    colortheme: "beaver"
    slide_level: 2
classoption: "aspectratio=169, t"
latex-auto-install: true
header-includes:
  - \setlength{\topsep}{0pt}
  - \setlength{\partopsep}{0pt}
  - \def\begincols{\begin{columns}}
  - \def\begincol{\begin{column}}
  - \def\endcol{\end{column}}
  - \def\endcols{\end{columns}}
  - \AtBeginSubsection{}
  - \AtBeginSection{}
  - \usepackage{xcolor}
  - \usepackage{booktabs}
  - \usepackage{multirow}
urlcolor: blue
---

<!-- cache: true -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE)
library(reticulate)

use_condaenv("/Users/amynussbaum/Library/r-miniconda-arm64/envs/r-reticulate/bin/python")

library(connectapi)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(lubridate)
library(stringr)
library(tidyr)

as_of_date <- today()
days_back <- 1000
top_n <- 30
report_from <- as_of_date - ddays(days_back)

client <- connect(server = "https://posit.ds.uchicago.edu/",
                  api_key = "XraWPYODS53ND3Y0Mw72coXKfnlQqmqj")

content_info <- connectapi::get_content(client) |>
  hoist(owner, owner_username = "username") |>
  select(guid, name, title, owner_username, dashboard_url)

usage_shiny <- get_usage_shiny(client,
                               from = report_from,
                               to = as_of_date,
                               limit = Inf
)

usage_non_shiny <- get_usage_static(client,
                                    from = report_from,
                                    to = as_of_date,
                                    limit = Inf
) |>
  rename(started = time) |>
  mutate(ended = started)

```

# Introduction

## Abstract

> Lab materials provide students opportunities for practice in new domains. However, existing tools for implementing interactive data science labs require installing software, which distracts from core lessons. We leverage R’s instructional ecosystem for browser-based, self-guided Python tutorials for data science courses. These tutorials require minimal setup, ideal for use across different classroom modalities. R’s framework can be successfully adapted for Python labs, lowering barriers to entry and strengthening student learning of core data science concepts.

## Motivation 



## Tools 



## Example

## Research Questions

* *RQ1*: Can frameworks developed for R-based courses be adapted for teaching data science in Python? 
* *RQ2*: Do browser-based, interactive labs improve understanding of data science concepts? 
* *RQ3*: Do such labs reduce barriers for students from non-coding backgrounds and increase confidence and self-efficacy?

# RQ1

## RQ1

Can frameworks developed for R-based courses be adapted for teaching data science in Python? 

> Yes! With some caveats--but mostly yes.

## Statistics by Class and Section

```{r, echo = FALSE}
quarter <- c("Winter 2025", "", "Spring 2025", "", "Sept. 2025", "Autumn 2025", "Winter 2026", "", "", "Spring 2026", "", "", "Total")
class <- c("119", "119", "119", "119", "119", "119", "119", "119", "120", "119", "119", "120", 2)
section <- c("02", "03", "01", "02", "01", "01", "01", "02", "01", "01", "02", "01", 12)
students <- c(34, 33, 22, 39, 12, 38, 19, 44, 29, 41, 37, 25, "373*")
format <- c("Required In-Class", "Asynchronous", "Asynchronous", "Required, In-Class", "Asynchronous", 
            "Asynchronous/Optional Lab", "Required Lab", "Required Lab", "Various",  "Required Lab", "Required Lab", "Various", "")

section_stats <- cbind(quarter, class, section, students, format)

kable(section_stats, format = "latex", escape = FALSE, booktabs = TRUE, 
      col.names = c("Quarter", "Class", "Section", "Students", "Format"), 
      align = c("r", "c", "c", "c", "c"), linesep = "") |>
  row_spec(c(2, 4, 5, 6, 9, 12), hline_after = TRUE, extra_latex_after = "%") |>
  row_spec(c(0, 13), bold = TRUE) |>
  kable_styling(position = "center", font_size = 8)
```

## DATA119 Tutorials

```{r, echo = FALSE}
DATA119 <- c("Data 119 Lab 3.5 - VIF and Categorical Variables", "Data 119 Lab 7 - Clustering",                     
             "Data 119 Lab 4 - Logistic Regression", "Data 119 Lab 2 - Correlation and SLR",            
             "Data 119 Lab 3 - Assumption Checking and MLR", "Data 119 Lab 6 - kNN and Confusion Matrices",     
             "Data 119 Lab 1 - Data Wrangling", "Data 119 Lab 8 - SQL",                            
             "Data 119 Lab 5 - Regularization", "Data 119 Lab 9 - Review")

usage <- bind_rows(usage_shiny, usage_non_shiny) |>
  mutate(quarter = ifelse(as.Date(started) < as.Date("03-15-2025", format = "%m-%d-%Y"), "Winter 2025", 
                          ifelse(as.Date(started) < as.Date("06-07-2025", format = "%m-%d-%Y"), "Spring 2025", 
                                 ifelse(as.Date(started) < as.Date("09-12-2025", format = "%m-%d-%Y"), "Sept. 2025", 
                                        ifelse(as.Date(started) < as.Date("12-13-2025", format = "%m-%d-%Y"), "Autumn 2025",
                                               ifelse(as.Date(started) < as.Date("03-14-2026", format = "%m-%d-%Y"), "Winter 2026", "Spring 2026"))))), 
         quarter_start = ifelse(as.Date(started) < as.Date("03-15-2025", format = "%m-%d-%Y"), as.Date("01-06-2025", format = "%m-%d-%Y"), 
                          ifelse(as.Date(started) < as.Date("06-07-2025", format = "%m-%d-%Y"), as.Date("03-24-2025", format = "%m-%d-%Y"), 
                                 ifelse(as.Date(started) < as.Date("09-12-2025", format = "%m-%d-%Y"), as.Date("08-25-2025", format = "%m-%d-%Y"), 
                                        ifelse(as.Date(started) < as.Date("12-13-2025", format = "%m-%d-%Y"), as.Date("09-29-2025", format = "%m-%d-%Y"),
                                               ifelse(as.Date(started) < as.Date("03-14-2026", format = "%m-%d-%Y"), as.Date("01-05-2026", format = "%m-%d-%Y"), 
                                                      as.Date("03-23-2026", format = "%m-%d-%Y"))))))) |>
  mutate(time_spent = (ymd_hms(ended) - ymd_hms(started))) |>
  mutate(quarter_day = as.Date(started) - as.Date(quarter_start)) |>
  filter(time_spent > 300) |>
  #select(content_guid, user_guid, time_spent) |>
  group_by(content_guid) |>
  mutate(avg_time = mean(time_spent), med_time = median(time_spent), 
         Q1 = quantile(time_spent, probs = c(0.25)), 
         Q3 = quantile(time_spent, probs = c(0.75))) |>
  mutate(IQR = Q3 - Q1) |>
  mutate(max_time = Q3 + 1.5*IQR) |>
  filter(time_spent < max_time) |>
  left_join(content_info, by = c(content_guid = "guid")) |>
  # if title is NA then substitute with name
  mutate(title = coalesce(title, name))
  
usage_119 <- usage |>  
  filter(title %in% DATA119)

usage_byl_119 <- usage_119 |>
  group_by(title) |>
  summarise(tot_usage = n(), med_time = round(median(as.numeric(time_spent/60)))) |> 
  #mutate(title = paste0(str_split(title, pattern = " ")[c(3, 4)], collapse = " "))
  mutate(title = sapply(str_split(title, pattern = " "), function(x) paste(x[3], x[4])))

usage_byl_119$topics <- c("Data wrangling, summary statistics, data visualization", 
                          "More data visualization, correlation, simple linear regression, some assumption checking",
                          "More assumption checking, multiple linear regression, checking significance", 
                          "Categorical variables, multicollinearity, variance inflation factor", 
                          "Logistic regression, including new model fit statistics like the confusion matrix", 
                          "Cross validation, ridge regression, LASSO", 
                          "$k$-nearest neighbors classification and regression", 
                          "$K$-means and hierarchical clustering", 
                          "Structured Query Language, including joins", 
                          "DATA119 Quarter Review")
```

```{r, echo = FALSE}
centerText <- function(text){
    paste0("\\multirow{2}{*}[0pt]{", text, "}")
  }

colnames(usage_byl_119) <- c(centerText("Lab"), centerText("Sessions"), "Median Time (m)", centerText("Topics"))

kable(usage_byl_119, format = "latex", escape = FALSE, booktabs = TRUE, 
      #col.names = , 
      align = c("l", "r", "c", "l"), linesep = "\\addlinespace") |>
  kable_styling(position = "center", font_size = 7) |>
  row_spec(0, bold = TRUE) |>
  column_spec(1, width = "1.2cm") |>
  column_spec(2, width = "0.9cm") |>
  column_spec(3, width = "1.5cm") |>
  column_spec(4, width = "8cm")
```

## DATA119: Usage by Time

```{r, echo = FALSE, fig.height = 4, fig.width=8, fig.align = "center"}
ggplot(usage_119, aes(x = time_spent/60, y = title)) + 
  geom_boxplot(color = "#800000") + 
  geom_vline(xintercept = 50, color = "gray20") + 
  scale_x_continuous("Time Spent (minutes)", 
                     breaks = seq(from = 0, to = 150, by = 30)) + 
  scale_y_discrete("", limits = rev, 
                   labels = c("Lab 1", "Lab 2", "Lab 3", "Lab 3.5", "Lab 4",
                              "Lab 5", "Lab 6", "Lab 7", "Lab 8", "Lab 9")) + 
  theme(axis.text.y = element_text(hjust = -0.14)) #+ 
  #ggtitle("DATA119 Tutorial Usage by Time")
```

## Usage for Lab 2: Correlation and Simple Linear Regression


## DATA120 Tutorials

```{r, echo = FALSE}
DATA120 <- c("Documentation Lab", "Debugging Lab", "Command Line Interface Lab",
             "Style Guide Lab", "Data 120 String Parsing Lab", "Sequences of Data Lab",
             "Functions and Raising Error Messages", "Trees Lab",                                       
             "Regular Expressions")

usage_120 <- usage |>
  filter(title %in% DATA120) |>
  mutate(title = factor(title, levels = DATA120))
```

```{r, echo = FALSE}
usage_byl_120 <- usage_120 |>
  group_by(title) |>
  summarise(tot_usage = n(), med_time = round(median(as.numeric(time_spent/60)))) |>
  mutate(title = c("Documentation*", "Debugging*", "Command Line Interface",
                   "Style Guide", "String Parsing", "Sequences of Data",
                   "Writing Functions", "Trees", "Regular Expressions"))

# usage_byl_120[1,1] <- "Documentation Lab"
# usage_byl_120[1,2] <- "Debugging Lab"

usage_byl_120$topics <- c("Reading and writing documentation for Python (review)", 
                          "Basic error types and debugging Python scripts (review)", 
                          "Using the command line interface to navigate through computers and run commands and scripts", 
                          "Motivating and adhering to a style guide for code", 
                          "String logic and methods with relevant applications", 
                          "Lists, tuples, dictionaries, and sets in Python", 
                          "Functions logic and keywords, raising error messages", 
                          "Trees as a data structure, some recursive functions",
                          "Regular expressions logic and operators, Python commands in the \\texttt{re} module")
```

```{r, echo = FALSE}
colnames(usage_byl_120) <- c(centerText("Lab"), centerText("Sessions"), "Median Time (m)", centerText("Topics"))

kable(usage_byl_120, format = "latex", escape = FALSE, booktabs = TRUE, 
      #col.names = c("Lab", "Sessions", "Median Time (m)", "Topics"), 
      align = c("l", "r", "c", "l"), linesep = "\\addlinespace") |>
  kable_styling(position = "center", font_size = 7.5) |>
  row_spec(0, bold = TRUE) |>
  column_spec(1, width = "1.9cm") |>
  column_spec(2, width = "0.9cm") |>
  column_spec(3, width = "1.5cm") |>
  column_spec(4, width = "7cm")
```

## DATA120: Usage by Time

```{r, echo = FALSE, fig.height = 4, fig.width=8, fig.align = "center"}
ggplot(usage_120, aes(x = time_spent/60, y = title)) + 
  geom_boxplot(color = c(rep("#275D38", 7), rep("#9CAF88", 2))) + 
  geom_vline(xintercept = 50, color = "gray20") + 
  scale_x_continuous("Time Spent (minutes)", 
                     breaks = seq(from = 0, to = 150, by = 30)) + 
  scale_y_discrete("", limits = rev, 
                   labels = c("Regular Expressions", "Trees", "Writing Functions",
                              "Sequences of Data", "String Parsing", "Style Guide", 
                              "Command Line Interface", "Debugging*", "Documentation*")) #+ 
  #theme(axis.text.y = element_text(hjust = -0.14)) + 
  #ggtitle("DATA120 Tutorial Usage by Time")

```

<!-- ## required labs ? -->

# RQ2

## RQ2

# RQ3

## RQ3

# Future Work

## Lessons Learned 1

Server performance matters!

>

Fixes and Future Plans

* Individual fine-tuning on each tutorial to improve performance
*  Reconfiguring server settings to clear up memory
* Migration to a new, larger, Data Science Institute-owned server over the summer. 

## Lessons Learned 2

* Recruitment is hard!
  + Two sections of DATA119 in Winter 2026 with 44 and 19 students
  + Ultimately, only 9 participants completed the pre- and post-quarter surveys. 
  
Fixes and Future Plans

* Repeated in-class visits by study recruiter
* Canvas announcements with links
  
## Future plans


